---
title: "Homework 4"
author: "Tony Daza"
date: "2022-10-20"
output: html_document
---


Here we go again. Let the coding begin.

  *Note, I accidentally named the project homework 4 instead of lab 4. I am sorry.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Step 1:
 - Load the required packages

This week we are focusing on data wrangling.

"Dplyr" is the package that we will use for data wrangling functions

"Janitor" helps with cleaning data, it may not be needed, but good to have.

"Tidyverse" includes ggplot2 as well as other packages and functions such as piping that are used throughout the code.

"Here" helps with the initial importing of data.

"Readr" is helpful for reading in data but Rio's import function is preferred for it's easy of use to import different types of files.

"Lahman" is a package with the data set we are using with all the baseball teams. The data set is called Teams

```{r, echo = TRUE}
library(needs)
needs(tidyverse, dplyr, janitor, here, readr, rio, Lahman)
view(Teams)

```

I am going to use the Minnesota Twins since I moved to Oregon from Minnesota. First, I will create a separate data set with only the twins data for year, wins, losses, runs scored, and runs allowed (Corresponding to name, yearID, W, L, R, RA)

```{r, echo = TRUE}
Twins <- Teams %>%
  filter(name == "Minnesota Twins", yearID >= 1980) %>%
  select(name, yearID, W, L, R, RA)
Twins

```


Cool! It worked!

Now we need to create a new variable that is the win percentage for the twins. 

If you are a twins fan you may want to skip this section

```{r, echo = TRUE}
Twins <- Twins %>%
  mutate(pwin = (W/(W+L))) %>%
  arrange(pwin)
Twins

```
Above is a list of the record of Twins seasons by year in ascending order by win percentage. 2016 was a rough year.

But let's look at the bright side and see the best years first by ordering the years in descending order by win percentage.

```{r, echo = TRUE}
Twins %>%
  arrange(desc(pwin))

```

Wow 2019 was a good year for them!

Well that was fun. Let's see the mean and standard deviation for win percentage

```{r, echo = TRUE}
Twins %>%
  summarise(mean(pwin),sd(pwin))

```
It appears that the mean win percentage is about 48% with a standard deviation of about 7.14%. I am not sure if that is good or not, but now I know.


That was child's play. Now we look at the full dataset.

```{r, echo = TRUE}
Teams <- Teams %>%
  select(name, yearID, W, L) %>%
  mutate(pwin = (W/(W+L))) %>%
  arrange(desc(pwin))
  
  

head(Teams, n = 10)

```
Above you can see the top ten seasons by win percentage over the years. 

1875 was an unforgettable year for the Boston Red Stockings.


Finally, to round out our baseball data. Below you will find a graph of the Detroit Tigers, The New York Yankees, and the San Diego Padres win percentage over the years.

```{r, echo= TRUE}
team_graph <- Teams %>%
  filter(name == "Detroit Tigers" | name == "New York Yankees" | name == "San Diego Padres", yearID >= 1900) %>%
  select(name, yearID, pwin) %>%
  ggplot(aes(x = yearID, y = pwin)) +
  geom_line(aes(color = name))+
  theme_minimal() +
  labs(x = "year_id", y = "w_pct", color = "Team")
team_graph


```
 
 
  *This does look slightly different than the example as there is an extra year or two of more recent data. I can remove those data points if needed

I hope you enjoyed our dive into baseball. We will now move on to talking about red blood cell counts and sports


```{r, echo = TRUE}
needs(readxl, readr)
ais <- read_excel(here("~/Desktop/Homework 4/data/ais.xlsx"), skip = 15)
head(ais)

```

Above is a peak at the AIS data set.We will come back to this one in a second.

```{r, echo = TRUE}
star <- import(here("Data/star.csv"))
head(star, n = 10)

```
Here is a look at the Star Data set

```{r, echo = TRUE}
sesame_st <- import(here("Data/sesame13.sav")) %>%
  janitor::clean_names()
head(sesame_st, n = 10)

```
Finally above is a peak at the sesame 13 data set.


And now back to the blood cells. Below is the average red blood cell count and average BMI by sport.

```{r, echo = TRUE}
AIS <- ais %>%
  select(rcc, bmi, sport) %>%
  group_by(sport) %>%
  summarise(avg_rcc = mean(rcc), 
            avg_bmi = mean(bmi))
AIS

export(AIS, "AIS_sport.xlsx")

export(AIS, "AIS_sport.sav")

```

Moving on. Let's look at the Sesame data and see if the average age of females was higher in school or at home.

```{r, echo = TRUE}
sesame_st %>%
  characterize("setting") %>%
  group_by(setting) %>%
  summarise(mean_age = mean(age))
  

```
The home group had a lower average age in months compared to the school group.


Finally let's look at the Star data
First we will look at the data using str() which will tell us that there are 12 variables (columns) and 5748 observations (rows)


```{r, echo = TRUE}
needs(psych)
str(star)

```

But personally I like the describe function from the psych package because it also gives the standard deviation and mean plus some other descriptive statistics.


```{r, echo = TRUE}
describe(star)
```

The next step is to remove the outliers in this case data that is 3 standard deviations above or below the mean

```{r, echo = TRUE}
star %>%
  characterize() %>%
  select(tmathss) %>%
  mutate(lower = (mean(tmathss) - sd(tmathss, na.rm = TRUE)*3)) %>%
  arrange(tmathss) %>%
  head(star, n = 5)

star %>%
  characterize() %>%
  select(tmathss) %>%
  mutate(upper = (mean(tmathss) + sd(tmathss, na.rm = TRUE)*3)) %>%
  arrange(desc(tmathss)) %>%
  head(star, n = 5)
```
These pieces show the lowest 5 and highest 5 scores as well as the cut off for 3 standard deviations above or below the mean.

```{r, echo = TRUE}
star %>%
  summarise(sd_low = (mean(tmathss) - (sd(tmathss)*3)),
         sd_high = (mean(tmathss) + (sd(tmathss*3))))
```
Here we can see the cut offs for the scores

```{r, echo = TRUE}
star_norm <- star %>%
  filter(tmathss < 628.9626, 
         tmathss > 342.3335)
head(star_norm)

```

That was a bit more challenging, but after some computations the values for 3 standard deviations above for the math score was 628.96 and 3 standard deviations below for math score was 342. 

After filtering for outliers, 5 observations were removed from the lower end of the data set.

Now for the more challenging part. I will standardized the math and reading scores. 

```{r, echo = TRUE}
star_stand <- star %>%
  select(tmathss, treadss) %>%
  mutate(tmathss = (tmathss - mean(tmathss))/sd(tmathss),
         treadss = (treadss - mean(treadss))/sd(treadss))
head(star_stand)

star_stand_means <- star_stand %>%
  summarise(mean(tmathss),
            sd(tmathss),
            mean(treadss),
            sd(treadss)) %>%
  round(digits = 3)
star_stand_means
         

```

Looks like it worked though there was a rounding error that prevented the mean from showing as zero, but the means of the standardized scores are essentially zero and with a little rounding it shows as zero.

```{r, echo = TRUE}
star_stand <- star %>%
  select(tmathss, treadss, sex, frl) %>%
  mutate(tmathss = (tmathss - mean(tmathss))/sd(tmathss),
         treadss = (treadss - mean(treadss))/sd(treadss))

star_stand_means_sex_frl <- star_stand %>%
  characterize() %>%
  group_by(sex, frl) %>%
  summarise(mean(tmathss),
            sd(tmathss),
            mean(treadss),
            sd(treadss))
star_stand_means_sex_frl

```

Above is a summary of the mean and sd of the standardized math and reading scores by sex and free and reduced price lunch status.

Based on this it looks like boys who do not receive frp lunch had  math and read scores higher than boys who did receive frp lunch. 

The difference was even more pronounced for girls who received frp lunch and girls who did not. With girls without frp lunch having higher math and reading scores on average.

There appears to be an interaction by gender with frl status being larger for girls than boys but I am not sure if it is significant without running further analyses.



At this point you can stop reading but I am curious if I can try to examine the data a little further.

```{r, echo = TRUE}
needs(MBESS, lm.beta)
cor(x = star_stand$tmathss, y = star_stand$treadss)
#Correlation between math and reading scores is 0.71 which is expected that they are strongly correlated in a positive direction

star_stand %>%
  ggplot(aes(x = treadss, color = frl)) +
  geom_bar(position = "dodge") + 
  facet_wrap(~sex)+
  theme_minimal()+
  labs(title = "Reading scores by Sex with FRP", x= "Standardized reading score", y = "Count", color = "FRP Lunch Status")+
  scale_color_viridis_d(option = "plasma")
#Standardized Reading scores by FRP Lunch status and sex distribution

star_stand %>%
  ggplot(aes(x = tmathss, color = frl)) +
  geom_bar(position = "dodge") + 
  facet_wrap(~sex)+
  theme_minimal()+
  labs(title = "Math scores by Sex with FRP", x= "Standardized Math score", y = "Count", color = "FRP Lunch Status")+
  scale_color_viridis_d()
#Standardized Math scores by FRP Lunch Status and sex distribution



```

It is clear from these graphs that there is a difference between the distributions of reading and math scores for students receiving FRP lunch and a smaller differences between boys and girls.

Perhaps if I have more time I will examine the exact differences.
