---
title: "Homework 5"
author: "Tony Daza"
date: "2022-10-27"
output: html_document
highlight: espresso

---
# Question 1
The YAML:
Change the highlight to Espresso


Homework 5, which is all about R markdown.


# Question 2
The set up:
You can't see it but it is there

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      eval = TRUE,
                      fig.height = 8.5,
                      fig.width = 6.5)
library(needs)
needs(tidyverse, rio, dplyr, here, readr, janitor)

```

# Question 3
Importing Data and Creating graphs:
Data from Ecls K Sample


Looking at t2m scale vs t1m scale by race and school type
t1mscale = T1 Math IRT scale score
t2mscale = T2 Math IRT scale score


```{r, }
eclsk <- import(here("Data/ecls-k_samp.sav")) %>%
  janitor::clean_names()

eclsk %>%
  select('ethnic')%>%
  distinct(ethnic) %>%
  arrange(ethnic) %>%
  characterize('ethnic')

# 1 is White, Non-Hispanic
# 2 is Black, Non-Hispanic
# 3 is Hispanic, Race Specified
# 4 is Hispanic, Race Not Specified
# 5 is Asian
# 6 is Native Hawaiian, Other Pacific Islander
# 7 is American Indian or Alaska Native
# 8 is More than one race, Non Hispanic

eclsk <- eclsk %>%
  characterize('school_type')

# 0 is Public
# 1 is Private

eclsk %>%
  select(t1mscale, t2mscale, school_type, ethnic) %>%
  filter(ethnic == "BLACK OR AFRICAN AMERICAN, NON-HISPANIC"| 
           ethnic == "HISPANIC, RACE NOT SPECIFIED" | 
           ethnic == "WHITE, NON-HISPANIC" | 
           ethnic == "HISPANIC, RACE SPECIFIED") %>%
  ggplot(aes(x = t1mscale, y = t2mscale), na.rm = TRUE) +
  geom_point(color = "black") +
  geom_smooth(aes(color = school_type), method = "lm") +
  facet_wrap(~ethnic) +
  theme_minimal()

```


As you can see, there are more white, non hispanic students than minority students in this sample and there scores tended to be higher than the Hispanic and Black students.


# Question 4
Mean and Standard Deviation: 
Mean & STD for T2Mscale

```{r}
t1mean <- mean(eclsk$t1mscale, na.rm = TRUE)
t1sd <- sd(eclsk$t1mscale, na.rm = TRUE)

t2mean <- mean(eclsk$t2mscale, na.rm = TRUE)
t2sd <- sd(eclsk$t2mscale, na.rm = TRUE)


```

# Question 5
Mean and Standard Deviation Continued:

T1 Mean/Std and T2 Mean/Std  


T1 Mean score = `r t1mean`  

T1 standard deviation score = `r t1sd`  

T2 Mean score = `r t1mean`  

T2 standard deviation score = `r t2sd`  

Difference in scores = `r t2mean - t1mean`  

*i.e. the average gain*


# Question 6
Loading Data Tutorial:


So let's say you are *new* to **R studio** and need to **load** in some data. *Follow along* to find out how to load in your data.

**Step 1**
Save data to your drive or computer
*Remember where you saved your data*

**Step 2**
Load **rio** to your console

\*You can load a package using the library( ) function

\# You will have to install rio first if it is not yet added

**Step 3**
Use the **import ()** function in rio to read in your data

\*when reading in the data you will need to copy the file pathway exactly and place it in quotes

**Step 4**
Save in your data using a nickname

You will need to use the function **<-** to rename a data set

**Step 5**
Look if your data was saved by calling the new name of the data set


For example: 
## pretend our data set is called Halloween and it is a csv file

library(rio)
Hallwn <- import("Halloween.csv")



```{r, echo=TRUE, eval = FALSE}

library(rio)
penguins <- import("Data/Penguins.csv")
penguins


```

In this example the Penguins data is saved in a folder called "Data"
Hence the pathway to the dataset is "Data/Penguins.csv"
