---
title: "Joining Data"
author: "Tony Daza"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


How to join data sets:  

Binding:
- bind_rows  
+ What we will use most often  

- bind_cols  
+ Can be dangerous if data rows are not in the same order  

2 types  
- Mutating join  
- Filter join  

Keys:
- For example Student ID  
+ Let's us know if data are lined up  
It may be more than one varaible across multiple columns  

Types of Keys:  
- Primary keys  
  + Used in your data set  
- Foreign Keys  
  + Found in your data set but not used for your dataset  
  
\* For .sav files the characterize( ) function makes the numeric labels characters if that information is present in the data  

Verifying the key:  
Use the count() function to ensure there is one key value for every row. There should not be repeats  

We can order them in descending order or filter for values of n greater then 1 will allow us to search for duplicates  

We can also use this in combination.
``` {r}
#count(year, percentile) %>%
#filter(n > 1)

```


What if we still can't find a key?  
- We can always use the row ID as the key if we know each entry is a unique data point  
  + This is called a surrogate key  
  
rowid_to_column() function in tidyverse will take the row id and create an id column  

#Joining

Types of Joining:  
- left_join  
  + Keep all the data on the left of dataset, drop non-matching (I.e. the dataset on the left of the function)  
- right_join  
  + Keep all data on the right of dataset, drop non-matching  
  
- inner_join  
  + Keeps only the matching data  
- full_join  
  + Keeps all of the data and adds NA values where data is missing  
  
\# case_when() can be helpful for recoding but can be tedious  


#Summary Data

Adding summary data into the dataset  





