---
title: "lab_1"
author: "Tony Daza"
date: "2022-10-07"
output: html_document
---
This is my Homework 1 for Intro to Data Science by Tony Daza

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(needs)
needs(tidyverse, rio, here, janitor, haven, psych, dplyr)

sesame13 <- import(here("Data/sesame13.sav")) %>%
  janitor::clean_names()

star <- read_csv(here("Data/star.csv")) %>% janitor::clean_names()

```

Apparently we also want to look at ggplot2 quickly. The mpg data set is pre-loaded in r. Below is a scatter plot of highway mpg by "displ" and another scatter plot of cylinders and city mpg

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + geom_point(col="gray70")
```

Now before I can make another scatter plot I need to check what the variables are in this data set.

```{r}
head(mpg)
```

And now another scatter plot of cylinders and city mpg

```{r}
ggplot(mpg, aes(x = cyl, y = cty)) + geom_point(col = "gray70")

```

Cool beans. Now we will continue with these data sets we imported.

This is the first 15 rows of the diamonds data set

```{r}
head(diamonds, 15)
```

This is the last 6 rows of the diamonds data set

```{r}

tail(diamonds)
```

This is the dimensions of the diamonds data set.

```{r}
str(diamonds)
```

But personally I like this view better which you get using the describe function from the psych package.

```{r}
describe(diamonds)
```

Here is a look at the different color types of diamonds.

```{r}
diamonds %>%
  count(diamonds$color)
```

Here is the code but without the piping.

```{r}
count(diamonds, color)
```

The END!